Assessing the Replicability of Psychological Science Through Pedagogy

### 2016-08-24 paper/paper.Rmd
fixes from student comments

author-note:
    "Thanks to the Stanford Department of Psychology and the Vice Provost for Undergraduate Education for funding to support the class."

@@

<!-- latex table generated in R 3.3.0 by xtable 1.8-2 package
 Fri Jun 24 12:02:11 2016 -->
\begin{table}[ht]
\footnotesize
\centering
\begin{tabular}{lllllrr}
  \hline
Citation & Expt & Data & Materials & Turk? & N (orig) & N (rep) \\ 
  \hline
Storm \& Stone (2015) & 3 & No & No & No &  48 &  61 \\ 
Lewis \& Oyserman (2015) & 4 & No & Yes & Yes & 122 & 128 \\ 
Scopelliti, Loewenstein, \& Vosgerau (2015) & 3 & No & Yes & Yes & 550 & 124 \\ 
Liverence \& Scholl (2015) & 1 & No & Some & No &  18 &  \\ 
Wang et al. (2015) & 2 & No & No & No & 219 & 397 \\ 
Sofer et al. (2015) & 1 & No & Yes & No &  48 &  95 \\ 
Ko, Sadler, \& Galinksy (2015) & 2 & Yes & Some & No &  40 &  40 \\ 
Atir, Rosenzweig, \& Dunning (2015) & 1b & No & Yes & Yes & 202 &  50 \\ 
Proudfoot, Kay, \& Koval (2015) & 1 & No & No & Yes &  80 &  84 \\ 
Zaval, Markowitz, \& Weber (2015) & 1 & Yes & Yes & Yes & 312 & 328 \\ 
Xu \& Franconeri (2015) & 1a & No & No & No &  12 &  27 \\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary of attempted replications.} 
\label{tab:summary}
\end{table}

@@

Once requested changes were made, the student conducted "Pilot B," using a handful of naive participants recruited from Mechanical Turk. The goal of this pilot was to ensure that participants were able to complete the experiment and did not report any substantial issues with instructions or technical details of the experiment. (All students were instructed to give participants a way to leave free-form comments at the end of the experiment but prior to debriefing text). At the conclusion of Pilot B, both the instructor and a TA reviewed the student's analytic code and its outputs on the data for all confirmatory analysis. The goal of this code review was to ensure that all planned analyses were specified correctly and with sufficient detail to permit inclusion in the broader analysis.

@@

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.
^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from pilot A sessions.] 
In 
`r sum(project_plan$power_standard == "80% power")`
cases, we powered the replication attempt to 80\% post-hoc power. In 
`r sum(project_plan$power_standard == "Original")` 
cases, we used the original sample size. In one case, the original experiment was apparently extremely over-powered in the post-hoc analysis, so we chose a smaller sample size that still would apparently yield high power.  In one case, we reduced the number of trials as part of the adaptation to mechanical turk, and in one case of a multi-step procedure where the first step consisted of stimulus generation, we decreased the N in the first stage of the procedure (details available in SOM).

@@

Students of the class and the instructors gave independent “replication ratings” to assess subjective judgement of the replication. Criterion is based on the theoretical interpretation of the original finding. (0 = non-replication, .5 is substantial hints of the same pattern or a significant but different pattern, 1 = full replication). We find that the these ratings of the student carrying out the replication and the instructors were in agreement for `r sum(d$replicated_TA == d$replicated_student,na.rm=T)` out of the `r nrow(d)` replications.  The average rating given by instructors and students were `r round(mean(d$replicated_TA, na.rm=TRUE), digits = 2)` and `r round(mean(d$replicated_student, na.rm=TRUE), digits = 2)` respectively.

@@

\begin{table}[ht]
\centering
\begin{tabular}{llrrr|r}
  \hline
Mediator & Value & No & Partial & Full & Total\\ 
  \hline
Replicated? &  &   2 &   6 &   3 & 11 \\ 
\hline
  Original on turk & Yes &   1 &   2 &   2 & 5 \\ 
   & No &   1 &   4 &   1 & 6\\ 
\hline
  High fidelity replication? & Yes &   2 &   3 &   1  & 6\\ 
   & No &  &   3 &   2 & 5\\ 
\hline
  Open materials? & Yes &   1 &   3 &   1  & 4\\ 
   & Some &  &   1 &   1 & 2\\ 
   &  No &   1 &   2 &   1 & 4\\ 
\hline
\end{tabular}
\caption{\label{tab:results} Counts of instructor team replication ratings, split by various different mediating variables.}
\end{table}

@@

The distribution of ratings are given in Table \ref{tab:results}, along with breakdowns by various factors of interest. Since our sample only includes 11 studies, we describe general trends here rather than attempting to conduct statistical tests (which would be dramatically under-powered after correcting for multiple comparisons). Overall, projects had a somewhat larger probability of being judged successful if it was judged to be closer to the original:
`r 100 * round(mean(d$replicated_TA[d$high_fidelity]), digits=2)`\% 
vs. 
`r 100 * round(mean(d$replicated_TA[d$high_fidelity == FALSE]), digits=2)`\%.
There were small numerical effects of being run on Mechanical Turk (
`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "Yes"]), digits=2)`\%
vs. 
`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "No"]), digits=2)`\%. 
) and having open materials (
`r 100 *round(mean(d$replicated_TA[d$open_materials!="No"]), digits=2)`\% 
vs. 
`r 100 *round(mean(d$replicated_TA[d$open_materials=="No"]), digits=2)`\%), though we do not believe these differences are interpretable given our sample size.

@@

Next, following @etz2016, we compared Bayes Factors between the original and replication. While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent discussions of the utility and validity of interpretating confidence intervals in this manner [@morey2016], as well as discrepencies in reporting effect sizes across the articles we included (which made computing confidence intervals difficult in some cases).

@@

This paper investigated motivations to save for retirement, and specifically whether changing participants' relationships to the future (making it feel more immanent) would cause them to save more. Our target was Study 4, which found that seeing the time to retirement in days (10,950) rather than years (30) caused participants to say they would begin saving sooner, even controlling for age, education, and income in a linear model (there was also a manipulation of whether saving was incremental that did not result in an effect). We failed to find a significant relationship between the time metric manipulation and saving time. However, when we did not include demographic controls, we did find a marginally-significant relationship. In exploratory analyses, we also observed an un-predicted effect of income such that participants who reported higher income selected to start saving sooner.

>>>>

author-note:
    "Thanks to the Stanford Department of Psychology and the Vice Provost for Graduate Education for funding to support the class."

@@

\newcommand{\specialcell}[2][c]{%
  \begin{tabular}[#1]{@{}c@{}}#2\end{tabular}}

<!-- latex table generated in R 3.3.0 by xtable 1.8-2 package
 Fri Jun 24 12:02:11 2016 -->
\begin{table}[ht]
\footnotesize
\centering
\begin{tabular}{lllllrr}
  \hline
Citation & Expt & \specialcell{Open\\Data?} & \specialcell{Open\\Materials?} & \specialcell{Original\\on Turk?} & N (orig) & N (rep) \\ 
  \hline
Storm \& Stone (2015) & 3 & No & No & No &  48 &  61 \\ 
Lewis \& Oyserman (2015) & 4 & No & Yes & Yes & 122 & 128 \\ 
Scopelliti, Loewenstein, \& Vosgerau (2015) & 3 & No & Yes & Yes & 550 & 124 \\ 
Liverence \& Scholl (2015) & 1 & No & Some & No &  18 &  19 \\ 
Wang et al. (2015) & 2 & No & No & No & 219 & 397 \\ 
Sofer et al. (2015) & 1 & No & Yes & No &  48 &  95 \\ 
Ko, Sadler, \& Galinksy (2015) & 2 & Yes & Some & No &  40 &  40 \\ 
Atir, Rosenzweig, \& Dunning (2015) & 1b & No & Yes & Yes & 202 &  50 \\ 
Proudfoot, Kay, \& Koval (2015) & 1 & No & No & Yes &  80 &  84 \\ 
Zaval, Markowitz, \& Weber (2015) & 1 & Yes & Yes & Yes & 312 & 328 \\ 
Xu \& Franconeri (2015) & 1a & No & No & No &  12 &  27 \\ 
   \hline
\end{tabular}
\caption{\label{tab:summary} Summary of papers and experiments that were targets for replication in our projects.} 
\label{tab:summary}
\end{table}

@@

Once requested changes were made, the student conducted "Pilot B," using a handful of naive participants recruited from Mechanical Turk. The goal of this pilot was to ensure that participants were able to complete the experiment and did not report any substantial issues with instructions or technical details of the experiment. (All students were instructed to give participants a way to leave free-form comments at the end of the experiment but prior to debriefing text). At the conclusion of Pilot B, both the instructor and a TA reviewed the student's analytic code and its outputs on the data for all confirmatory analyses. The goal of this code review was to ensure that all planned analyses were specified correctly and with sufficient detail to permit inclusion in the broader analysis.

@@

We eventually used a hybrid, case-by-case selection criterion that attempted to maximize the success of the project while staying within the constraints of our budget.
^[We initially were allocated \$1,500 but one student contributed personal research funds, leading to a total cost of $\approx$ \$1,700 for all studies. We attempted to set payment for our studies at approximately \$6/hour based on timing estimated from pilot A sessions.] 
In 
`r sum(project_plan$power_standard == "80% power")`
cases, we powered the replication attempt to 80\% post-hoc power. In 
`r sum(project_plan$power_standard == "Original")` 
cases, we used the original sample size. In one case, the original experiment was  extremely over-powered in the post-hoc analysis, so we chose a smaller sample size that still would yield high power.  In one case, we reduced the number of trials as part of the adaptation to Mechanical Turk, and in one case of a multi-step procedure where the first step consisted of stimulus generation, we decreased the N in the first stage of the procedure (details available in SOM).

@@

Students of the class and the instructors gave independent “replication ratings” to assess subjective judgement of the replication. Our rating system was based on the theoretical interpretation of the original finding, with 0 = no support for the original interpretation, .5 = some support for the original interpretation, 1 = full replication). We found that the ratings of the student carrying out the replication and the instructors were in agreement for `r sum(d$replicated_TA == d$replicated_student,na.rm=T)` out of the `r nrow(d)` replications.  The average rating given by instructors and students were `r round(mean(d$replicated_TA, na.rm=TRUE), digits = 2)` and `r round(mean(d$replicated_student, na.rm=TRUE), digits = 2)` respectively.

@@

\begin{table}[ht]
\centering
\begin{tabular}{rllllllll}
  \hline
\multicolumn{2}{c}{Replicated} & \multicolumn{2}{c}{Original on Turk?} & \multicolumn{2}{c}{High fidelity?} & \multicolumn{3}{c}{Open Materials?} \\
  \hline
   &  & No & Yes & Yes & No & No & Some & Yes \\ 
  No &  2 &  1 &  1 &  2 & 0 &  1 & 0 &  1 \\ 
  Partial & 6 & 4 & 2 & 4 & 2 & 2 & 1 & 3 \\ 
  Full & 3 & 1 & 2 & 1 & 2 & 1 & 1 & 1 \\ 
  \hline
  Total & 11 & 6 & 5 & 5 & 6 & 4 & 2 & 5 \\
   \hline
\end{tabular}
\caption{\label{tab:results} Counts of instructor team replication ratings, split by mediating variables.}
\end{table}

<!-- \begin{table}[ht] -->
<!-- \centering -->
<!-- \begin{tabular}{llrrr|r} -->
<!--   \hline -->
<!-- Mediator & Value & No & Partial & Full & Total\\  -->
<!--   \hline -->
<!-- Replicated? &  &   2 &   6 &   3 & 11 \\  -->
<!-- \hline -->
<!--   Original on turk & Yes &   1 &   2 &   2 & 5 \\  -->
<!--    & No &   1 &   4 &   1 & 6\\  -->
<!-- \hline -->
<!--   High fidelity replication? & Yes &   2 &   3 &   1  & 6\\  -->
<!--    & No &  &   3 &   2 & 5\\  -->
<!-- \hline -->
<!--   Open materials? & Yes &   1 &   3 &   1  & 4\\  -->
<!--    & Some &  &   1 &   1 & 2\\  -->
<!--    &  No &   1 &   2 &   1 & 4\\  -->
<!-- \hline -->
<!-- \end{tabular} -->
<!-- \caption{\label{tab:results} Counts of instructor team replication ratings, split by various different mediating variables.} -->
<!-- \end{table} -->

@@

The distribution of ratings are given in Table \ref{tab:results}, along with breakdowns by various factors of interest. Since our sample only includes 11 studies, we describe general trends here rather than attempting to conduct statistical tests (which would be dramatically under-powered after correcting for multiple comparisons). Overall, projects had a somewhat larger probability of being judged successful if they were judged to be closer to the original:
`r 100 * round(mean(d$replicated_TA[d$high_fidelity]), digits=2)`\% 
vs. 
`r 100 * round(mean(d$replicated_TA[d$high_fidelity == FALSE]), digits=2)`\%.
There were small numerical effects of being run on Mechanical Turk (
`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "Yes"]), digits=2)`\%
vs. 
`r 100 *round(mean(d$replicated_TA[d$original_on_turk == "No"]), digits=2)`\%. 
) and having open materials (
`r 100 *round(mean(d$replicated_TA[d$open_materials!="No"]), digits=2)`\% 
vs. 
`r 100 *round(mean(d$replicated_TA[d$open_materials=="No"]), digits=2)`\%), though we do not believe these differences are interpretable given our sample size.

@@

Next, following @etz2016, we compared Bayes Factors between the original and replication. While we preregistered an analysis determining whether the original reported effect size is included in the 95\% confidence interval found in the replication, we instead opted for a straightforward comparison of point estimates. This choice was motivated by recent discussions of the utility and validity of interpreting confidence intervals in this manner [@morey2016], as well as discrepencies in reporting effect sizes across the articles we included (which made computing confidence intervals difficult in some cases).

@@

This paper investigated motivations to save for retirement, and specifically whether changing participants' relationships to the future (making it feel more imminent) would lead them to report that they would begin saving for retirement sooner. Our target was Study 4, which found that seeing the time to retirement in days (10,950) rather than years (30) caused participants to say they would begin saving sooner, even controlling for age, education, and income in a linear model (there was also a manipulation of whether saving was incremental that did not result in an effect). We failed to find a significant relationship between the time metric manipulation and saving time. However, when we did not include demographic controls, we did find a marginally-significant relationship. In exploratory analyses, we also observed an un-predicted effect of income such that participants who reported higher income selected to start saving sooner.
